---
title: "Entropy Exercises"
output: html_document
date: "2025-04-05"
---

## load your packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(tidyverse, cmdstanr, brms, posterior,bayesplot)

register_knitr_engine(override = FALSE)


```

### Lets start by simulating some response times that are dependent on the probabilities from the psychometric function. Here lets assume that the mean response times follow:

$$
\mu_{RT} = RT_{int} + RT_{\beta} \cdot (- p  \cdot log(p) - (1-p) \cdot log(1-p))
$$

```{r}
## Define parameters


# Simulate the stimulus sequence:

# function for the psychometric function:


# generate probabilities

# Generate the mean response times from the formula above:

# plot stimulus sequences vs probabilities:

# plot stimulus sequences vs mean response times:


```



### As we don't observe the probabilities we don't observe the mean response times. We observe a noisy measure of them

### Thus how do we go from our mean response times to the noisy response times that we observe?


```{r}
# Covert each mean response time to a response time at trial t using a distribution.


# plot stimulus sequences vs response times.


```



## Code up a stanmodel that takes the stimulus sequences, binary responses and response times and returns the parameters you choose above!

```{cmdstan, output.var="model"}

// Data block for the input sequence and binary responses
data {

  // Define the number of datapoints this is defined by an integer

  // Define a vector that is as long as the number of datapoints of real numbers (the stimulus sequence)  

  // Define a vector that is as long as the number of datapoints of real numbers (the response times)  
  
  // Define an array that is as long as the number of datapoints of integers (the binary responses)

    
}

// This is the parameters block. Here we define the free parameters.
parameters {
  // Define the threshold
  
  // Define the slope
  
  // Define the lapse rate

  // Define the RT_int

  // Define the RT_beta

  // Define the RT_sigma
  
  
}

// This is the model block here we define how we believe the model is and the priors of the parameters.

model {
  // priors
  
  // threshold prior
  
  // slope prior
  
  // lapse rate prior
  
  // RT_int prior
  
  // RT_beta prior
  
  // RT_sigma rate prior
  
  
  // the model and the likelihood for both measures (one for the response times and one for the binary choices)
    

    
}
```



### Now lets see if the model runs:

```{r}
# if you wrote a seperate stan-file load it here:


# Otherwise you can access the model from the above chuck like this:

# model$sample()

# and then you have to input the sampler parameters, number of chains and the data in a list


```

### Checking covergence:

```{r}

```


### Check estimates compared to the simulated:


```{r}

```

