---
title: "Pathfinder exercises"
output: html_document
date: "2025-04-05"
---

## load your packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(tidyverse, cmdstanr, brms, posterior,bayesplot)

register_knitr_engine(override = FALSE)

```



## Using Pathfinder

### Check the following link and try to sample the same model using pathfinder:
https://mc-stan.org/cmdstanr/reference/model-method-pathfinder.html

### What are the estimates from pathfinder compared to the standard approach?

```{r}

```


### Now try and setup an experimental setting where each additional stimulus value gets choosen based on estimates from pathfinder.

### perhaps select the next stimulus value at the estimated mean threshold.

```{r}

```

### How much faster can pathfinder do this compared to the full MCMC-sampling of stan?

```{r}

```


### How do you select the next stimulus value if you want to minimize the posterior spread of the threshold?

```{r}

```


### What about the slope or the lapse rate?

```{r}
```


### try and plot the trajectory of the estimate of the parameters across trials for different ways of selecting stimulus values.

```{r}

```

