---
title: "Psychometric Exercises"
output: html_document
date: "2025-04-05"
---

## load your packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(tidyverse, cmdstanr, brms, posterior,bayesplot)

register_knitr_engine(override = FALSE)


```

### Simulate a subject's psychometric function for a stimulus sequence from -20 to 20 with:

$$
\alpha = 5, 
\beta = 2,
\lambda = 0.05
$$

```{r}
# define parameters 
alpha <- 5
beta <- 2
lambda <- 0.05
# Simulate the stimulus sequence:
stims <- seq(-20,20,0.1)

# function for the psychometric function:
Psychometric_function <-function(alpha, beta, lambda, x=stims, gamma){
  sim <- gamma +(1-gamma-lambda)*pnorm(x, alpha, beta)
  return(sim)
}

# generate probabilities

# plot stimulus sequences vs probabilities:

```


### In these experiments we don't observe the probability of the our subjects responding we observe the response. 

### Thus how do we go from our probabilities to the actual binary outcomes?


```{r}
# Covert each probability to a binary response by using a distribution.


# plot stimulus sequences vs probabilities with the realized binary responses.


```


## Code up a stanmodel that takes the stimulus sequences, binary responses that assumes the same psychometric function as you simulated data with. You can do this in the code-chunk below or make a seperate stan-file

```{cmdstan, output.var="model"}

// Data block for the input sequence and binary responses
data {

  // Define the number of datapoints this is defined by an integer

  // Define a vector that is as long as the number of datapoints of real numbers (the stimulus sequence)  

  // Define an array that is as long as the number of datapoints of integers (the binary responses)

    
}

// This is the parameters block. Here we define the free parameters.
parameters {
  // Define the threshold
  
  // Define the slope
  
  // Define the lapse rate
  
  
}

// This is the model block here we define how we believe the model is and the priors of the parameters.

model {
  // priors
  
  // threshold prior
  
  // slope prior
  
  // lapse rate prior
  
  
  
  // the model and the likelihood:
    

    
}
```


### Now lets see if the model runs:

```{r}
# if you wrote a seperate stan-file load it here:


# Otherwise you can access the model from the above chuck like this:
# model$sample()
# and then you have to input the sampler parameters, number of chains and the data in a list



```

### Checking covergence:

```{r}

```


### Check estimates compared to the simulated:


```{r}

```

