---
title: "Pathfinder Hint"
output: html_document
date: "2025-04-05"
---

## load your packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(tidyverse, cmdstanr, brms, posterior,bayesplot)

register_knitr_engine(override = FALSE)
```



## Using Pathfinder

### Check the following link and try to sample the same model using pathfinder:
https://mc-stan.org/cmdstanr/reference/model-method-pathfinder.html

### What are the estimates from pathfinder compared to the standard approach?

```{r}
# use pathfinder to estimate the same as above:
model$pathfinder(data = list(x = x, y = y, N = length(x)))

# then extract the estimates and plot them against the full mcmc model.

```


### Now try and setup an experimental setting where each additional stimulus value gets choosen based on estimates from pathfinder.

### perhaps select the next stimulus value at the estimated mean threshold.

```{r}
# start with an inital stimulus value say 0 then try to run the pathfinder model using this.

# after you get the pathfinder estimates select a new stimulus value using those estimates (you decide where to place them!). 

# Then generate the next stimulus value and generate a binary response for that stimulus value and run the model again (now on 2 datapoints).

# repeat for how many trials you like (loop over trials!).

```

### How much faster can pathfinder do this compared to the full MCMC-sampling of stan?

```{r}

#  keep track of estimation time and comparing to the full MCMC sampling method. 

# One can extract the timne taken to fit with fit$time()

```


### How do you select the next stimulus value if you want to minimize the posterior spread of the threshold?

```{r}

# play around with the decision rule for selecting the next stimulus. Use the information from pathfinder to guide your decision

```


### What about the slope or the lapse rate?

```{r}

# play around with the decision rule for selecting the next stimulus. Use the information from pathfinder to guide your decision

```


### try and plot the trajectory of the estimate of the parameters across trials for different ways of selecting stimulus values.

```{r}

# make a plot with iteration of the experiment (experiment trial) on the x-axis and the estimated parameter (from pathfinder on the y-axis).

# Extra points for including the uncertainty of the estimated parameters.

```

