---
title: "DDM Hint"
output: html_document
date: "2025-04-05"
---

## load your packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(tidyverse, cmdstanr, brms, posterior,bayesplot,RWiener)

register_knitr_engine(override = FALSE)
```

## Lets start by simulating some data from the DDM. This can be done using the RWiener package. 
## Find the random number generator from the package and generate 100 samples and plot these

```{r}
# the function is called rweiner and needs the following parameters: 
# n number of samples, alpha = bounary seperation, tau = non-decision-time, beta = bias, delta = drift-rate

# define parameters 
n = 100

boundary =
  
bias = 
  
ndt = 
  
drift = 

rwiener()

```


## Write a stan model that fits this data:

## Check the slides!

```{cmdstan, output.var="model"}

// Data block for the input sequence and binary responses
data {

  // Define the number of datapoints this is defined by an integer

  // Define a vector that is as long as the number of datapoints of real numbers (the response times)  

  // Define an array that is as long as the number of datapoints of integers (the binary responses)

    
}

// This is the parameters block. Here we define the free parameters.
parameters {
  // Define the boundary
  
  // Define the bias
  
  // Define the non-decision-time
  
  // Define the Drift-rate
  
  
}

// This is the model block here we define how we believe the model is and the priors of the parameters.

model {
  // priors
  
  // boundary prior
  
  // bias prior
  
  // non-decision-time prior
  
  // drift-rate prior
  
  
  // The model and the likelihood:
    

}
```

## Fit the model
```{r}
fit = model$sample(iter_warmup = 500,
                   iter_sampling = 500,
                   adapt_delta = 0.9,
                   parallel_chains = 4,
                   chains = 4,
                   data = list(resp = resp, RT = RT, N = length(x)))
```


## check covergence of the model and the estimated parameters compared to the simulated
```{r}
# investigation of divergences max-treedepths, effective sample size and Rhat.

# traceplots

# pairplots!
```




# Now lets infuse the DDM with the psychometric function. Or atleast the stimulus intensity

## One apporach is to make the drift-rate vary as a function of the probability p.

$$
\delta_t = p \cdot (1-p) \cdot \delta
$$

## Use the psychometric simulations to generate the probabilities (p) and then generate Response times and binary responses from the
## DDM / RWiener distribution 


```{r}

boundary =
  
bias = 
  
ndt = 
  
drift = 
  
alpha = 
  
beta = 
  
lambda = 
  
x = 
  
p = 

delta_t = 

df = rwiener()

```


## Now plot the binary responses together with the response times. 

```{r}


```



# write another stan program that uses the above equation and data:

```{cmdstan, output.var="model"}

// Data block for the input sequence and binary responses
data {

  // Define the number of datapoints this is defined by an integer

  // Define a vector that is as long as the number of datapoints of real numbers (the response times)  

  // Define a vector that is as long as the number of datapoints of real numbers (the stimuli)  

  
  // Define an array that is as long as the number of datapoints of integers (the binary responses)

    
}

// This is the parameters block. Here we define the free parameters.
parameters {
  
  // Define the threshold
  
  // Define the slope
  
  // Define the lapse rate
  
  // Define the boundary
  
  // Define the bias
  
  // Define the non-decision-time
  
  // Define the Drift-rate
  
  
}

// This is the model block here we define how we believe the model is and the priors of the parameters.

model {
  // priors
  
  // boundary prior
  
  // bias prior
  
  // non-decision-time prior
  
  // drift-rate prior
  
  // threshold prior
  
  // slope prior
  
  // lapse rate prior
  
  
  
  // The model and the likelihood:
    

}
```


# check covergence of the model and the estimated parameters compared to the simulated
```{r}
# investigation of divergences max-treedepths, effective sample size and Rhat.


```


### Check estimates compared to the simulated:

```{r}
## take the draws from the fit using the posterior package's function as_draws_df() and plot the posterior distribution of each parameter together with some indication of what the simulated value was.


# extracting the draws from the model fit:

draws = as_draws_df(fit$draws(c("beta","alpha","lambda","boundary","bias","ndt","drift")))
```


