---
title: "PSI Hint"
output: html_document
date: "2025-04-05"
---

## load your packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(tidyverse, cmdstanr, brms, posterior,bayesplot)

register_knitr_engine(override = FALSE)
```

### Simulate a subject's psychometric function for a stimulus sequence from -20 to 20 with:

$$
\alpha = 5, 
\beta = 2,
\lambda = 0.05
$$

```{r}
# define parameters 
alpha =
  
beta = 
  
lambda = 
# Simulate the stimulus sequence:
x = seq(-20,20,by = 0.1)
  
# function for the psychometric function:
psychometric = function(x,alpha,beta,lambda){
  # Another tip is that one can use the pnorm() function or the error function in the pracma package.
  
}

# generate probabilities

p = psychometric(x,alpha,beta,lambda)

# plot stimulus sequences vs probabilities:

## make a dataframe with x and p and use ggplot to plot it as points or a line.

```


### In these experiments we don't observe the probability of the our subjects responding we observe the response. 

### Thus how do we go from our probabilities to the actual binary outcomes?


```{r}
# Covert each probability to a binary response by using a distribution.

## the distribution is the bernoulli distribution, which is the simplified version of the binominal distribution


# plot stimulus sequences vs probabilities with the realized binary responses.

## make a dataframe with x and p and the binary responses from above and use ggplot to plot it as points or a line.

```


## Code up a stanmodel that takes the stimulus sequences, binary responses that assumes the same psychometric function as you simulated data with. You can do this in the code-chunk below or make a seperate stan-file

```{cmdstan, output.var="model"}

// Data block for the input sequence and binary responses
data {

  // Define the number of datapoints this is defined by an integer

  // Define a vector that is as long as the number of datapoints of real numbers (the stimulus sequence)  

  // Define an array that is as long as the number of datapoints of integers (the binary responses)

    
}

// This is the parameters block. Here we define the free parameters.
parameters {
  // Define the threshold
  
  // Define the slope
  
  // Define the lapse rate
  
  // remember constraints on the parameters!
  
}

// This is the model block here we define how we believe the model is and the priors of the parameters.

model {
  // priors
  
  // threshold prior
  
  // slope prior
  
  // lapse rate prior
  
  
  // Remember constraints on the parameters, either one relies on the definition of the parameters or use prior-distributions that only allow    // the right domain!
  
  
  
  // the model and the likelihood:
    
    
  // bernoulli() is a function in stan as well as erf() and Phi().

    
}
```


### Now lets see if the model runs:

```{r}
# if you wrote a seperate stan-file load it here:


# Otherwise you can access the model from the above chuck like this:
# model$sample()
# and then you have to input the sampler parameters, number of chains and the data in a list

fit = model$sample(iter_warmup = 200,
                   iter_sampling = 200,
                   adapt_delta = 0.9,
                   parallel_chains = 4,
                   chains = 4,
                   data = list(x = x, y = y, N = length(x))
                   )

```

### Checking covergence:
```{r}
# investigation of divergences max-treedepths, effective sample size and Rhat.

# traceplots


# pairplots!
```


### Check estimates compared to the simulated:

```{r}
## take the draws from the fit using the posterior package's function as_draws_df() and plot the posterior distribution of each parameter together with some indication of what the simulated value was.


```

