---
title: "Entropy Hints"
output: html_document
date: "2025-04-05"
---

## load your packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(tidyverse, cmdstanr, brms, posterior,bayesplot)

register_knitr_engine(override = FALSE)


```

### Lets start by simulating some response times that are dependent on the probabilities from the psychometric function. Here lets assume that the mean response times follow:

$$
\mu_{RT} = RT_{int} + RT_{\beta} \cdot (- p  \cdot log(p) - (1-p) \cdot log(1-p))
$$

```{r}
## Define parameters
alpha
  
beta
  
lambda 

rt_int

rt_beta 


# Simulate the stimulus sequence:

x = seq(-20,20,by = 0.1)
  
# function for the psychometric function:
psychometric = function(x,alpha,beta,lambda){
  # Another tip is that one can use the pnorm() function or the error function in the pracma package.
  lambda + (1-2*lambda) * pnorm((x-alpha)/sqrt(2)*beta,0,1)
  
}

# generate probabilities
p = psychometric(x,alpha,beta,lambda)

# Generate the mean response times from the formula above:

#The entropy can be written as: (-p*log(p) - (1-p) * log(1-p))



# plot stimulus sequences vs probabilities:


# plot stimulus sequences vs mean response times:


```


### As we don't observe the probabilities we don't observe the mean response times. We observe a noisy measure of them

### Thus how do we go from our mean response times to the noisy response times that we observe?


```{r}
# Covert each mean response time to a response time at trial t using a distribution.

# One might use a normal distribution which then requires another parameter (the standard deviation or residual variance here.)
rt_sd =


# plot stimulus sequences vs response times.

  
  
binary_resp = rbinom(length(p),1,p)
```



## Code up a stanmodel that takes the stimulus sequences, binary responses and response times and returns the parameters you choose above!

```{cmdstan, output.var="model"}


// Data block for the input sequence and binary responses
data {

  // Define the number of datapoints this is defined by an integer
  
  int N;
  
  // Define a vector that is as long as the number of datapoints of real numbers (the stimulus sequence)  
  vector[N] x;

  // Define a vector that is as long as the number of datapoints of real numbers (the response times)  
  vector[N] RT;


  // Define an array that is as long as the number of datapoints of integers (the binary responses)
  array[N] int binary_resp;



}

// This is the parameters block. Here we define the free parameters.
parameters {
  // Define the threshold
  real alpha;
  // Define the slope
  real<lower=0> beta;
  // Define the lapse rate
  real<lower=0, upper = 0.5> lambda;
  
  // Define the parameters for the response time part:
    
}

// This is the model block here we define how we believe the model is and the priors of the parameters.

model {
  // priors
  
  // threshold prior
  alpha ~ normal(0,10);
  // slope prior
  beta ~ normal(0,10);  
  // lapse rate prior
  lambda ~ normal(0,0.1);
  

  // Priors for the response time parameters
  
  

  // Remember constraints on the parameters, either one relies on the definition of the parameters or use prior-distributions that only allow    // the right domain!
  
  // the model and the likelihood:
    
  vector[N] p = lambda + (1-2*lambda) * Phi((x-alpha)/sqrt(2) * beta);
    
  // bernoulli() is a function in stan as well as erf() and Phi().
  binary_resp ~ bernoulli(p);
 
  // write out the likelihood for the response times using the same distribution as you used to simulate.
  
}

```



### Now lets see if the model runs:

```{r}
# if you wrote a seperate stan-file load it here:


# Otherwise you can access the model from the above chuck like this:

## add the data!

fit = model$sample(iter_warmup = 200,
                   iter_sampling = 200,
                   adapt_delta = 0.9,
                   parallel_chains = 4,
                   chains = 4,
                   data = list()
                   )

# and then you have to input the sampler parameters, number of chains and the data in a list


```

### Checking covergence:

```{r}
# investigation of divergences max-treedepths, effective sample size and Rhat.

# traceplots

# pairplots!

```


### Check estimates compared to the simulated:


```{r}
## take the draws from the fit using the posterior package's function as_draws_df() and plot the posterior distribution of each parameter together with some indication of what the simulated value was.

# extracting the draws from the model fit:
full_mcmc = as_draws_df(fit$draws(c("beta","alpha","lambda","rt_int","rt_beta","rt_sd")))


```


## What might be the problem with using a normal distribution / likelihood for the response time?

## Can you come up with another distribution or some transformation that could help with the problem?


